The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-22 manuscript.Rmd
little polish of the Research Compendia section text

Containers provide a high degree of isolation that is often desirable when attempting to capture a specific computational environment so that others can reproduce and extend a research result.
Many computationally intensive research projects depend on specific versions of original and third-party software packages in diverse languages, joined together to form a pipeline through which data flows.
New releases of even just a single piece of software in this pipeline can break the entire workflow, making it difficult to find the error and for others to reuse existing pipelines, and making the original the results irreproducible and not extandable.
The chance of a substantial disruption like this is high in a multi-year research project where key pieces of third-party software may have several major updates over the duration of the project.
The classical "paper" article is insufficient to adequately communicate the knowledge behind such research projects \citep[cf.][]{donoho_invitation_2010,marwick_how_2015}.
\citet{gentleman_statistical_2007} coined the term _Research Compendium_ for a dynamic document together with supporting data and code.
They use the R package system \cite{core_writing_1999} as for the functional prototype to structuring, validation, and distribution of research compendia.
This concept has been taken up and extended\footnote{See full literature list at \href{https://research-compendium.science/}{https://research-compendium.science/}.}, not the least by applying containerisation and other methods for managing computing environments---see Section&nbsp;\nameref{envs}.
Containers give the researcher an isolated environment to assemble these research pipelines with specific versions of software.
Research workflows in containers are safe from contamination from other activities occuring on the researcher's computer, for example the installation of newest version of packages for teaching demonstrations.
Given the users in this scenario, i.e., often academics with limited formal software development training, templates and assistance with containers around research compendia is essential.

In many fields we see that a typical unit of research for a container is a research report or journal article, where the container holds the compendium, or self-contained set of data (or connections to data elsewhere) and code files needed to fully reproduce the article \citep{marwick_packaging_2018}.
The package \pkg{rrtools} ([https://github.com/benmarwick/rrtools](https://github.com/benmarwick/rrtools)) provides a template and convienence functions to apply good practices for research compendia, including a starter `Dockerfile`.
Images of compendium containers can be hosted on services such as Docker&nbsp;Hub for convienient sharing among collaborators and others. 
Similarly, packages such as \pkg{containerit} and \pkg{dockerfiler} can be used to manage the `Dockerfile` to be archived with a compendium on a data repository (e.g. [Zenodo](https://zenodo.org/), [Dataverse](https://dataverse.org/), [Figshare](https://figshare.com/), [OSF](https://osf.io/)).
A typial compendium's `Dockerfile` will pull a rocker image fixed to a specific version of R, and install R packages from the MRAN repository to ensure the package versions are tied to a specific date, rather than the most recent version.
A future researchers can download then compendium from the repository and run the included `Dockerfile` build a new image that recreates the computational environment used to produce the original research results.
If building the image fails, the human-readable instructions in a `Dockerfile` are a the starting point for rebuilding the environment.
Further safeguarding practices are currently under development or not part of common practice yet, such as preservation of images \citep{emsley_framework_2018} and storing them alongside `Dockerfile`s \citep[cf.][]{nust_opening_2017}, or pinning of system libraries beyond the tagged base images, which may be seen as stable or dynamic depending on the applied time scale \citep[see discussion on `debian:testing` base image in][]{RJ-2017-065}.

@@

A recommendation of the recent National Academies' report on _Reproducibility and Replicability in Science_ is that journals "consider ways to ensure computational reproducibility for publications that make claims based on computations" \citep{NASEM2019}.
In fields such as Political Science and Economics, journals are increasingly adopting policies requiring authors to publish the code and data required to reproduce computational findings reported in published manuscripts, subject to independent verification \citep{Jacoby2017,Vilhuber2019,Alvarez2018,Christian2018,Eubank2016,King1995}.
Problems with the computational environment, installation and availability of software dependencies are common. R is gaining popularity in these communities (in a sample of 105 replication packages published by the _American Journal of Political Science_ (AJPS) over 65% use R). 
Projects such as NSF Whole Tale use the Rocker Project community images with the goal of improving the reproducibility of published research artifacts and simplifying the publication and verification process for both authors and journals by reducing errors and time spent specifying the environment.

@@

The ordering of authors following DN as first author is alphabetical.
DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
DE wrote the introduction and the section about Containerization and Rocker.
RC \& EH contributed to the section on interfaces for Docker in R.
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
BM wrote the section on research compendia and made the project Binder-ready.
HN \& JN co-wrote the section on enterprise production environments (\pkg{googleComputeEngineR}, Google Cloud Run, Kubernetes).
KR wrote the section about \pkg{wholepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
CW wrote the sections on Whole Tale and contributed the publication reproducibility audit use case.
NX wrote the section on \pkg{liftr}.
All authors approved the final version.
This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page}, \href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history}, and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide a detailed view on the respective contributions.

>>>>

Containers provide a high degree of isolation that is often desirable when attempting to capture a specific computational environment so that others can reproduce and extend a research result.
Many computationally intensive research projects depend on specific versions of original and third-party software packages in diverse languages, joined together to form a pipeline through which data flows.
New releases of even just a single piece of software in this pipeline can break the entire workflow, making it difficult to find the error and difficult for others to reuse existing pipelines.
These breakages can make the original the results irreproducible and not extandable.
The chance of a substantial disruption like this is high in a multi-year research project where key pieces of third-party software may have several major updates over the duration of the project.
The classical "paper" article is insufficient to adequately communicate the knowledge behind such research projects \citep[cf.][]{donoho_invitation_2010,marwick_how_2015}.
\citet{gentleman_statistical_2007} coined the term _Research Compendium_ for a dynamic document together with supporting data and code.
They use the R package system \cite{core_writing_1999} as the functional prototype for structuring, validation, and distribution of research compendia.
This concept has been taken up and extended\footnote{See full literature list at \href{https://research-compendium.science/}{https://research-compendium.science/}.}, not the least by applying containerisation and other methods for managing computing environments---see Section&nbsp;\nameref{envs}.
Containers give the researcher an isolated environment to assemble these research pipelines with specific versions of software.
Research workflows in containers are safe from contamination from other activities occuring on the researcher's computer, for example the installation of newest version of packages for teaching demonstrations.
Given the users in this scenario, i.e., often academics with limited formal software development training, templates and assistance with containers around research compendia is essential.

In many fields we see that a typical unit of research for a container is a research report or journal article, where the container holds the compendium, or self-contained set of data (or connections to data elsewhere) and code files needed to fully reproduce the article \citep{marwick_packaging_2018}.
The package \pkg{rrtools} ([https://github.com/benmarwick/rrtools](https://github.com/benmarwick/rrtools)) provides a template and convienence functions to apply good practices for research compendia, including a starter `Dockerfile`.
Images of compendium containers can be hosted on services such as Docker&nbsp;Hub for convienient sharing among collaborators and others. 
Similarly, packages such as \pkg{containerit} and \pkg{dockerfiler} can be used to manage the `Dockerfile` to be archived with a compendium on a data repository (e.g. [Zenodo](https://zenodo.org/), [Dataverse](https://dataverse.org/), [Figshare](https://figshare.com/), [OSF](https://osf.io/)).
A typial compendium's `Dockerfile` will pull a rocker image fixed to a specific version of R, and install R packages from the MRAN repository to ensure the package versions are tied to a specific date, rather than the most recent version.
Future researchers can download the compendium from the repository and run the included `Dockerfile` to build a new image that recreates the computational environment used to produce the original research results.
If building the image fails, the human-readable instructions in a `Dockerfile` are the starting point for rebuilding the environment.
Further safeguarding practices are currently under development or not part of common practice yet, such as preservation of images \citep{emsley_framework_2018} and storing them alongside `Dockerfile`s \citep[cf.][]{nust_opening_2017}, or pinning of system libraries beyond the tagged base images, which may be seen as stable or dynamic depending on the applied time scale \citep[see discussion on `debian:testing` base image in][]{RJ-2017-065}.

@@

A recommendation of the recent National Academies' report on _Reproducibility and Replicability in Science_ is that journals "consider ways to ensure computational reproducibility for publications that make claims based on computations" \citep{NASEM2019}.
In fields such as Political Science and Economics, journals are increasingly adopting policies requiring authors to publish the code and data required to reproduce computational findings reported in published manuscripts, subject to independent verification \citep{Jacoby2017,Vilhuber2019,Alvarez2018,Christian2018,Eubank2016,King1995}.
Problems with the computational environment, installation and availability of software dependencies are common. R is gaining popularity in these communities (in a sample of 105 replication packages published by the _American Journal of Political Science_ (AJPS) over 65% use R). 
The NSF-funded Whole Tale project uses the Rocker Project community images with the goal of improving the reproducibility of published research artifacts and simplifying the publication and verification process for both authors and journals by reducing errors and time spent specifying the environment.

@@

The ordering of authors following DN as first author is alphabetical.
DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
DE wrote the introduction and the section about Containerization and Rocker.
RC \& EH contributed to the section on interfaces for Docker in R.
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
BM wrote the section on research compendia and made the project Binder-ready.
HN \& JN co-wrote the section on enterprise production environments (\pkg{googleComputeEngineR}, Google Cloud Run, Kubernetes).
KR wrote the section about \pkg{holepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
CW wrote the sections on Whole Tale and contributed the publication reproducibility audit use case.
NX wrote the section on \pkg{liftr}.
All authors approved the final version.
This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page}, \href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history}, and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide a detailed view on the respective contributions.
